---
title: "Analysis of Reported Car Crashes in Montgomery County, MD"
authors: "Adriana Vance and Elena Koshkin"
format: html
code-fold: true
editor: visual
execute: 
  warning: false 
---

```{r, results = "hide", message = FALSE, warning = FALSE}
Crash <- read.csv("Crash.csv")

library(ggplot2)
library(dplyr)
library(tigris)
library(sf)

```

## Visualization 1

```{r, results = "hide", message = FALSE, warning = FALSE}
sf_data <- st_as_sf(Crash, coords = c("Longitude", "Latitude"), crs = 4326)

# Get the county boundaries for the entire state of Maryland
maryland_counties <- counties("MD")
# Filter Montgomery County
montgomery_county <- maryland_counties[maryland_counties$NAME == "Montgomery", ]

montgomery_roads <- roads("MD", "Montgomery")


filtered_crash_data <- sf_data %>%
  group_by(ACRS.Report.Type) %>%
  filter(ACRS.Report.Type == "Fatal Crash") %>%
  summarize(count = n())

montgomery_roads %>%
  ggplot() + 
   geom_sf(data = montgomery_county) +
    geom_sf(data = montgomery_roads, alpha = 0.3, color = "blue") +
  
  geom_sf(data = filtered_crash_data, aes(color = "Location of Crashes", shape = "Location of Crashes")) + 
  labs(
    title = "Fatal Crashes in Montgomery County", subtitle = "2015-2023", caption = "Source: https://catalog.data.gov/dataset/crash-reporting-drivers-data") + 
  theme_void() + 
  scale_color_manual(
    name = "",
    values = c("Location of Crashes" = "red")
  ) + 
  scale_shape_manual(
    name = "",
    values = c("Location of Crashes" = 8) 
  )
  
```

Our first visualization displays a map of all of the fatal crashes that have occurred in Montgomery County, MD in the past 8 years (2015-2023). This map displays the county boundaries as well as all of the roads in the county in blue. This shows that many crashes are concentrated in downtown areas like Rockville. It also indicates that many fatal crashes occur on major roads such as Interstate 270. In terms of policy implications, this could show policymakers that higher concentrations of cars and higher speeds are correlated with more fatal crashes. It is therefore important to enforce speed limits and other safety precautions in these high-risk areas.

## Visualization 2

```{r}

Crash %>%
  group_by(Collision.Type) %>%
  summarize(count = n()) %>%
  filter(count > 10000) %>%
  filter(Collision.Type != "OTHER") %>%
  ggplot() + 
  geom_col(mapping = aes(y = count, x = Collision.Type, fill = Collision.Type), alpha = 0.5) + 
  theme(axis.text.x = element_blank()) + 
  labs(
    x = "Collision Type",
    y = "Total Collisions",
    caption = "Source: https://catalog.data.gov/dataset/crash-reporting-drivers-data ", 
    fill = "Collision Type", title = "Most Common Collisions in Montgomery County", subtitle = "Totals from 2015-2023")

```

Our next visualization displays the five most common types of collisions in Montgomery County. The bar graph shows each type of collision from 2015-2023. The most obvious conclusion is that rear-ends are the most common type of collision in Montgomery County, MD. This may tell us something about the types of scenarios during which drivers tend to collide. Since the majority of rear-end collisions happen due to [distracted driving](https://www.daveabels.com/rear-end-collisions-frequent-type-collision/#:~:text=A%20report%20from%20the%20National%20Transportation%20Safety%20Board%20found%20that,stop%20to%20avoid%20a%20collision.), this may suggest that we should implement a policy that more strictly penalizes distracted driving, or run an awareness campaign about distracted driving.

## Visualization 3

```{r}

Crash %>%
  filter(Vehicle.Year > 1990 & Vehicle.Year < 2023) %>%
  filter(Injury.Severity == "FATAL INJURY") %>% 
  ggplot(aes(x= Vehicle.Year, fill = Injury.Severity)) +
  geom_histogram(alpha = 0.4, color = "#000000", fill = "#0200F9") + labs(
    title = "Fatal Injuries by Vehicle Year", subtitle = "Years 1940-2023",
    x = "Vehicle Year", 
    y = "Count", 
    caption = "Source: https://catalog.data.gov/dataset/crash-reporting-drivers-data") + 
  theme_light()
```
Our third visualization is a histogram showing the frequency of fatal injuries across vehicle years. We had originally hypothesized that the frequency of fatal injuries would be higher in older vehicle years, suggesting that older cars should be taken off the road. However, this histogram shows that there isn't a skewed distribution of fatal injuries across vehicle years and there are just as many fatal injuries occurring in newer cars than in older cars. This might suggest that vehicle safety standards can still stand to improve or that driver safety training has room for improvement, but that in the meantime, older vehicles do not need to be prohibited on the road.

## Visualization 4

```{r}

ggplot(Crash, aes(x=Speed.Limit, fill=Driver.At.Fault)) +
  geom_density(alpha=.4) +
  labs(
    x = "Speed Limit",
    y = "Density",
    caption = "Source: https://catalog.data.gov/dataset/crash-reporting-drivers-data", fill = "Driver at Fault", title = "Speed Limits by Driver at Fault", subtitle = "Across Crashes"
    ) + 
  theme(legend.title = element_text(colour="maroon", size=10, face="bold"))
```
This last visualization is a density plot of the speed limits of the roads where crashes occurred. The plot is separated is grouped into three categories depending on whether the driver was at fault for the crash. This visualization demonstrates that while most crashes, regardless of whether the driver was at fault or not, occur on roads where the speed limit is 35 mph, there is still a significant concentration of crashes on roads with speed limits less than 35 mph. The policy implication that could be derived from this graph is the necessity for increased driver safety instruction as part of the required training for obtaining a driver's license. Although most *fatal* crashes occur on major roads (visualization 1), this shows that the most crashes occur at slower speeds emphasizing the need for better safety instruction in more populated/residential areas. It is important that drivers are prioritizing road safety at 5 mph, just as much as at 40mph or above. 